Water Desalination Using Polyelectrolyte Hydrogel: Gibbs~Ensemble Modeling

### 2022-10-14 MDPI/proof/manuscript.v5/orig.tex
proofreading

%=================================================================
% MDPI internal commands
\firstpage{1} 
\makeatletter 
\setcounter{page}{\@firstpage} 
\makeatother
\pubvolume{1}
\issuenum{1}
\articlenumber{0}
\pubyear{2022}
\copyrightyear{2022}
\externaleditor{Academic Editor: \hl{Firstname Lastname}} %MDPI: please add the Academic Editor
\datereceived{9 August 2022} 
%\daterevised{} % Only for the journal Acoustics
\dateaccepted{10 October 2022} 
\datepublished{} 
%\datecorrected{} % Corrected papers include a "Corrected: XXX" date in the original paper.
%\dateretracted{} % Corrected papers include a "Retracted: XXX" date in the original paper.
\hreflink{https://doi.org/} % If needed use \linebreak
%\doinum{}
%------------------------------------------------------------------
% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
%\pdfoutput=1

@@

% Authors, for the paper (add full first names)
\Author{Mikhail Laktionov %MDPI: Please carefully check the accuracy of names and affiliations. 
%MDPI: The name of this author is different from the submitting system. Please confirm which one is correct.
 $^{1, 2}$\orcidA{}, Lucie Nová $^{1}$\orcidB{} and Oleg V. Rud $^{1, 3}$*\orcidC{}}

@@

% MDPI internal command: Authors, for citation in the left column
\AuthorCitation{Laktionov, M.; Nová, L.; Rud, O.V.}
% If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad Department of Physical and Macromolecular Chemistry, Faculty of Science, Charles University in Prague, Prague 12800, Czech Republic
; miklakt@gmail.com (M.L.); lucie.nova@natur.cuni.cz (L.N.)\\
$^{2}$ \quad Saint-Petersburg National Research University of Information Technologies, Mechanics and Optics, Saint-Petersburg 197101, Russia\\
$^{3}$ \quad Institute of Macromolecular Compounds of Russian Academy of Sciences, Saint-Petersburg 199004, Russia}

@@

We propose the desalination process as a cascade of gel compressions and decompressions, lowering the  supernatant salinity.
%it is worth to note that we are not limited with this process only.
In this process, the gel is supposed to be compressed/decompressed in an \emph{open system %MDPI: Please confirm if the italics should be retained.
%Oleg Rud: Confirm, let it retain
} and in a \emph{closed system}.
The compression in the \emph{open system} assumes that the gel is in thermodynamic equilibrium with a huge (effectively \textbf{infinite}) reservoir of aqueous solution \reffig{fig: open}, whereas 
the \emph{closed system} implies that the gel is in equilibrium with a \textbf{finite %MDPI: Please confirm if the bold should be retained.
%Oleg Rud: Confirm, let it retain
} reservoir of an aqueous solution \reffig{fig: closed}.

@@

	\caption{The hydrogel compressed in \emph{open} and in \emph{closed} system.
		Red and blue circles represent $\na$ and$\cl$ ions. $V_0$ is the volume which gel has in free swelling equilibrium state.\label{fig:open and closed}}
\end{figure}
%In the case of \emph{open system} the gel is the grand canonical ensemle, as $\mu, V, T = \mbox{Const}$, while the case of \emph{closed system} the gel and the explicit reservoir are a variation of Gibbs ensemble \cite{Panagiotopoulos1988b} performed with constant volume for the both boxes, equal chemical potentials of ion species at constant temperature.

@@

\includegraphics[width=0.7\textwidth]{{figures/simgibbs}.pdf}\caption{Diamond-like network in the simulation box. Color code represents
the individual ion types (red: $\na$, blue: $\cl$)
and the hydrogel (gray: neutral segment ($\AH$), cyan: charged segment
($\A$)). \label{fig:diamond}}
\end{figure}
Each monomer unit of the network carries a negative elementary electric charge.
Except for the gel monomers, the monovalent co- and counter-ions, $\cl$ and $\na$, are present in the simulation box.
The total electric charge of all the particles in the box is zero; therefore the number of $\na$ ions exceeds the number of $\cl$ ions by $\Ngel$.

@@

\begin{table}[H]
\caption{Estimates of  %MDPI: Please confirm if the format of the table should be retained, e.g., the color of the font, the text direction, vertical line
% Oleg Rud: Confirm the colors and the directions of the table must retain
desalination efficiency. All units are calculated per one mol of gel segments. Values in brackets correspond to crossection of `red' and `orange' lines on the plots in Figures \ref{fig: PV and CV} and \ref{fig: NV and CN}.
\label{tab: table}}

\begin{adjustwidth}{-\extralength}{0cm}

		\newcolumntype{C}{>{\centering\arraybackslash}X}
\begin{tabularx}{1\fulllength}{@{\extracolsep{\fill}}ll|lc|c|l|ll}
\noalign{\hrule height 1pt}
\multicolumn{1}{l|}{\textbf{\boldmath{$c_{s}^{0}$}, mM}} & \textbf{\boldmath{$c_{s}^{5}$}, mM} & \multicolumn{1}{l|}{\textbf{\boldmath{$n_{\cl}^{0}$}, mM}} & \textbf{\boldmath{$n_{\cl}^{5}$}, mM} & \textbf{\boldmath{$\Delta \Vgel$}, L} & \textbf{\boldmath{$|W$}\textbar , J/L} & \multicolumn{2}{c}{\textbf{\boldmath{$W^{id}$}, J/L}}\tabularnewline

\noalign{\hrule height 0.5pt} 
\multirow{2}{*}{{\small{}$91.47$}} & \multirow{2}{*}{} & \multicolumn{2}{l|}{{\small{}$57.08\pm0.122\enskip\longrightarrow$}} & \multirow{2}{*}{{\small{}2.74}} & \multirow{2}{*}{{\small{}$95.4\pm1.9$}} &  & \multirow{2}{*}{}\tabularnewline
 &  & \multicolumn{2}{r|}{{\small{}$71.75\pm0.024$}} &  &  & \multirow{8}{*}{{\small{}\hspace{-1em}}\textcolor{teal}{\small{}$\left.\begin{array}{l}
\\
\\
\\
\\
\\
\\
\\
\\
\end{array}\right\rbrace $$\rotatebox{90}{\hspace{-5em}{\color{teal}\ensuremath{\begin{array}{cc}
W^{id}= & 52.9\\
W^{\mathrm{sim}}= & 202.8\pm3.2\\
R_{w}= & 0.54
\end{array}}}}$}} & \tabularnewline
\cline{1-6} \cline{2-6} \cline{3-6} \cline{4-6} \cline{5-6} \cline{6-6} 
\multicolumn{2}{l|}{{\small{}$89.41\pm0.23\enskip\longrightarrow$}} & \multirow{2}{*}{{\small{}56.90}} & \multirow{2}{*}{} & \multirow{2}{*}{{\small{}2.72}} & \multirow{2}{*}{{\small{}$109.1\pm1.7$}} &  & \multirow{2}{*}{}\tabularnewline
\multicolumn{2}{r|}{{\small{}$73.63\pm0.03$}} &  &  &  &  &  & \tabularnewline
\cline{1-6} \cline{2-6} \cline{3-6} \cline{4-6} \cline{5-6} \cline{6-6} 
\multirow{2}{*}{\textcolor{teal}{\small{}63.93}} & \multirow{2}{*}{} & \multicolumn{2}{l|}{\textcolor{teal}{\small{}$37.28\pm0.08\enskip\longrightarrow$}} & \multirow{2}{*}{\textcolor{teal}{\small{}3.26}} & \multirow{2}{*}{\textcolor{teal}{\small{}$100.9\pm1.7$}} &  & \tabularnewline
 &  & \multicolumn{2}{r|}{\textcolor{teal}{\small{}$50.579\pm0.013$}} &  &  &  & \multirow{8}{*}{{\small{}\hspace{-1em}}\textcolor{olive}{\small{}$\left.\begin{array}{l}
\\
\\
\\
\\
\\
\\
\\
\\
\end{array}\right\rbrace $$\rotatebox{90}{\hspace{-5em}\ensuremath{\begin{array}{cc}
W^{id}= & 38.2\\
W^{\mathrm{sim}}= & 207.3\pm2.2\\
R_{w}= & 0.53
\end{array}}}$}}\tabularnewline
\cline{1-6} \cline{2-6} \cline{3-6} \cline{4-6} \cline{5-6} \cline{6-6} 
\multicolumn{2}{l|}{\textcolor{teal}{\small{}$62.05\pm0.15\enskip\longrightarrow$}} & \multirow{2}{*}{\textcolor{teal}{\small{}37.17}} & \multirow{2}{*}{} & \multirow{2}{*}{\textcolor{teal}{\small{}3.18}} & \multirow{2}{*}{\textcolor{teal}{\small{}$107.4\pm1.4$}} &  & \tabularnewline
\multicolumn{2}{r|}{\textcolor{teal}{\small{}$48.21\pm0.02$}} &  &  &  &  &  & \tabularnewline
\cline{1-6} \cline{2-6} \cline{3-6} \cline{4-6} \cline{5-6} \cline{6-6} 
\multirow{2}{*}{\textcolor{olive}{\small{}44.91}} & \multirow{2}{*}{} & \multicolumn{2}{l|}{\textcolor{olive}{\small{}$23.75\pm0.06\enskip\longrightarrow$}} & \multirow{2}{*}{\textcolor{olive}{\small{}3.82}} & \multirow{2}{*}{\textcolor{olive}{\small{}$106.7\pm1.5$}} &  & \tabularnewline
 &  & \multicolumn{2}{r|}{\textcolor{olive}{\small{}$35.911\pm0.010$}} &  &  & \multirow{8}{*}{{\small{}\hspace{-1em}}\textcolor{red}{\small{}$\left.\begin{array}{l}
\\
\\
\\
\\
\\
\\
\\
\\
\\
\end{array}\right\rbrace \rotatebox{90}{\hspace{-5em}\ensuremath{\begin{array}{cc}
W^{\mathrm{id}}= & 43.8\ {\color{red}{\scriptstyle (41.0)}}\\
W^{\mathrm{sim}}= & 222.3\pm2.5\\
R_{w}= & 0.52\ {\scriptstyle (0.46)}
\end{array}}}$}} & \tabularnewline
\cline{1-6} \cline{2-6} \cline{3-6} \cline{4-6} \cline{5-6} \cline{6-6} 
\multicolumn{2}{l|}{\textcolor{olive}{\small{}$43.46\pm0.12\enskip\longrightarrow$}} & \multirow{2}{*}{\textcolor{olive}{\small{}24.27}} & \multirow{2}{*}{} & \multirow{2}{*}{\textcolor{olive}{\small{}3.91}} & \multirow{2}{*}{\textcolor{olive}{\small{}$106.4\pm1.1$}} &  & \tabularnewline
\multicolumn{2}{r|}{\textcolor{olive}{\small{}$30.795\pm0.008$}} &  &  &  &  &  & \tabularnewline
\cline{1-6} \cline{2-6} \cline{3-6} \cline{4-6} \cline{5-6} \cline{6-6} 
\multirow{2}{*}{\textcolor{red}{\small{}31.93}} & \multirow{2}{*}{} & \multicolumn{2}{l|}{\textcolor{red}{\small{}$14.66\pm0.04\enskip\longrightarrow$}} & \multirow{2}{*}{\textcolor{red}{\small{}4.20}} & \multirow{2}{*}{\textcolor{red}{\small{}$107.9\pm1.4$}} &  & \tabularnewline
 &  & \multicolumn{2}{r|}{\textcolor{red}{\small{}$25.297\pm0.006$}} &  &  &  & \multirow{8}{*}{{\small{}\hspace{-1em}}\textcolor{orange}{\small{}$\left.\begin{array}{l}
\\
\\
\\
\\
\\
\\
\\
\\
\\
\end{array}\right\rbrace \rotatebox{90}{\hspace{-5em}\ensuremath{\begin{array}{cc}
W^{id}= & 66.6\ {\scriptstyle (22.0)}\\
W^{\mathrm{sim}}= & 218.7\pm2.2\\
R_{w}= & 0.53\ {\scriptstyle (0.49)}
\end{array}}}$}}\tabularnewline
\cline{1-6} \cline{2-6} \cline{3-6} \cline{4-6} \cline{5-6} \cline{6-6} 
\multicolumn{2}{l|}{\textcolor{red}{\small{}$30.03\pm0.08\enskip\longrightarrow$}} & \multirow{2}{*}{\textcolor{red}{\small{}14.55}} & \multirow{2}{*}{} & \multirow{2}{*}{\textcolor{red}{\small{}$\begin{array}{c}
4.17\\
{\scriptstyle (3.22)}
\end{array}$}} & \multirow{2}{*}{\textcolor{red}{\small{}$\begin{array}{c}
115.6\pm0.9\\
{\scriptstyle (57.3)}
\end{array}$}} &  & \tabularnewline
\multicolumn{2}{r|}{\textcolor{red}{\small{}$\begin{array}{c}
18.585\pm0.004\,\\
{\scriptstyle (22.32)}
\end{array}$}} &  &  &  &  &  & \tabularnewline
\cline{1-6} \cline{2-6} \cline{3-6} \cline{4-6} \cline{5-6} \cline{6-6} 
\multirow{2}{*}{\textcolor{orange}{\small{}22.30}} & \multirow{2}{*}{} & \multicolumn{2}{l|}{\textcolor{orange}{\small{}$8.84\pm0.03\enskip\longrightarrow$}} & \multirow{2}{*}{\textcolor{orange}{\small{}$\begin{array}{c}
4.75\\
{\scriptstyle (3.76)}
\end{array}$}} & \multirow{2}{*}{\textcolor{orange}{\small{}$\begin{array}{c}
108.1\pm1.3\\
{\scriptstyle (68.3)}
\end{array}$}} &  & \tabularnewline
 &  & \multicolumn{2}{r|}{\textcolor{orange}{\small{}$\begin{array}{c}
\text{17.945}\pm\text{0.004}\\
{\scriptstyle (14.56)}
\end{array}$}} &  &  & \multirow{7}{*}{{\small{}\hspace{-1em}}\textcolor{magenta}{\small{}$\left.\begin{array}{l}
\\
\\
\\
\\
\\
\\
\\
\end{array}\right\rbrace \rotatebox{90}{\hspace{-5em}\ensuremath{\begin{array}{cc}
W^{id}= & 69.7\\
W^{\mathrm{sim}}= & 227.5\pm2.3\\
R_{w}= & 0.55
\end{array}}}$}} & \\[1ex]
\cline{1-6} \cline{2-6} \cline{3-6} \cline{4-6} \cline{5-6} \cline{6-6} 
\multicolumn{2}{l|}{\textcolor{orange}{\small{}$20.74\pm0.06\enskip\longrightarrow$}} & \multirow{2}{*}{\textcolor{orange}{\small{}8.83}} & \multirow{2}{*}{} & \multirow{2}{*}{\textcolor{orange}{\small{}$4.71$}} & \multirow{2}{*}{\textcolor{orange}{\small{}$110.8\pm0.8$}} &  & \tabularnewline
\multicolumn{2}{r|}{\textcolor{orange}{\small{}$11.082\pm0.002$}} &  &  &  &  &  & \tabularnewline
\cline{1-6} \cline{2-6} \cline{3-6} \cline{4-6} \cline{5-6} \cline{6-6} 
\multirow{2}{*}{\textcolor{magenta}{\small{}10.90}} & \multirow{2}{*}{} & \multicolumn{2}{l|}{\textcolor{magenta}{\small{}$3.00\pm0.01\enskip\longrightarrow$}} & \multirow{2}{*}{\textcolor{magenta}{\small{}6.08}} & \multirow{2}{*}{\textcolor{magenta}{\small{}$106.9\pm1.1$}} &  & \tabularnewline
 &  & \multicolumn{2}{r|}{\textcolor{magenta}{\small{}9.011$\pm$0.002}} &  &  &  & \tabularnewline
\cline{1-6} \cline{2-6} \cline{3-6} \cline{4-6} \cline{5-6} \cline{6-6} 
\multicolumn{2}{l|}{\textcolor{magenta}{\small{}$9.83\pm0.05\enskip\longrightarrow$}} & \multirow{2}{*}{\textcolor{magenta}{\small{}3.12}} & \multirow{2}{*}{} & \multirow{2}{*}{\textcolor{magenta}{\small{}5.78}} & \multirow{2}{*}{\textcolor{magenta}{\small{}$119.4\pm1.0$}} &  & \tabularnewline
\multicolumn{2}{r|}{\textcolor{magenta}{\small{}$3.862\pm0.001$}} &  &  &  &  &  & \\[2ex]
\noalign{\hrule height 1pt}
\end{tabularx}
\end{adjustwidth}
\end{table}

@@

\section{Conclusions}
\textls[-5]{We have modeled compression of a polyelectrolyte gel in thermodynamic equilibrium with a supernatant aqueous solution of limited amount.
We have shown that compression of the gel decreases the supernatant salinity.
We employed this phenomenon to model water desalination.
The desalination was done as a sequential combination of two processes:
(1) swelling of the gel in an \emph{open system}, exchanging ions with a large reservoir at constant salinity;
(2) compression of the gel in a \emph{closed system}, during which the gel exchanges ions with a small reservoir, affecting its salinity.
We estimated the energy consumption needed for producing one liter of potable water from brine and have shown that the proposed gel compression method may compete with modern desalination~technologies. %MDPI: we noticed that the supplementary file has been uploaded in submitting system. please confirm if there is a supplementary, if so, please cite it in main text and add section of supplementary materils “\supplementary{The following are available at \linksupplementary{s1}, Figure S1: title, Table S1: title, Video S1: title.} ”. Please note that the reference in supplementary should be also included in mian text.
% Oleg Rud: There is no need of supplementary materials anymore, since we added the method description into the main text.
}

@@

%\institutionalreview{\hl{~}}%MDPI: In this section, please add the Institutional Review Board Statement and approval number for studies involving humans or animals. Please note that the Editorial Office might ask you for further information. Please add ``The study was conducted according to the guidelines of the Declaration of Helsinki, and approved by the Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE (protocol code XXX and date of approval).'' OR ``Ethical review and approval were waived for this study, due to REASON (please provide a detailed justification).'' OR ``Not applicable'' for studies not involving humans or animals. You might also choose to exclude this statement if the study did not involve humans or animals.

%\informedconsent{\hl{~}}%MDPI: Any research article describing a study involving humans should contain this statement. Please add ``Informed consent was obtained from all subjects involved in the study.'' OR ``Patient consent was waived due to REASON (please provide a detailed justification).'' OR ``Not applicable'' for studies not involving humans. You might also choose to exclude this statement if the study did not involve humans. Written informed consent for publication must be obtained from participating patients who can be identified (including by the patients themselves). Please state ``Written informed consent has been obtained from the patient(s) to publish this paper'' if applicable.

%\dataavailability{\hl{~}} %MDPI: In this section, please provide details regarding where data supporting reported results can be found, including links to publicly archived datasets analyzed or generated during the study. Please refer to suggested Data Availability Statements in section ``MDPI Research Data Policies'' at \url{https://www.mdpi.com/ethics}. You might choose to exclude this statement if the study did not report any data.

%\acknowledgments{\hl{This research was supported by the Czech Science Foundation (grant 19-17847Y) and the 
%government of the Russian Federation (grant number 14.W03.31.0022).} %MDPI: The contents in acknowledgments are the same as those in funding. please confirm if it can be deleted in acknowledgments
 % Confirm it can be deleted in acknowledjements
%}

@@

\bibitem[Miller(2003)]{Miller2003}
Miller, J.
\newblock \emph{Review of Water Resources and Desalination Technologies}; 
Sandia National Laboratories (SNL): Albuquerque, NM, USA, 2003;
\newblock Technical Report; 
\newblock https://doi.org/10.2172/809106; 
%MDPI: Please add the name of the publisher and the location of it.
.

\newblock {\em {Physical Chemistry}}, 9th ed.; Oxford University Press:
  New York, NY, USA, 2010.
\newblock
  https://doi.org/10.2172/809106.%}{\detokenize{10.2172/809106}}}.

>>>>

%=================================================================
% MDPI internal commands
\firstpage{1} 
\makeatletter 
\setcounter{page}{\@firstpage} 
\makeatother
\pubvolume{1}
\issuenum{1}
\articlenumber{0}
\pubyear{2022}
\copyrightyear{2022}
\externaleditor{Academic Editor: \hl{Firstname Lastname} %MDPI: please add the Academic Editor
}
\datereceived{9 August 2022} 
%\daterevised{} % Only for the journal Acoustics
\dateaccepted{10 October 2022} 
\datepublished{} 
%\datecorrected{} % Corrected papers include a "Corrected: XXX" date in the original paper.
%\dateretracted{} % Corrected papers include a "Retracted: XXX" date in the original paper.
\hreflink{https://doi.org/} % If needed use \linebreak
%\doinum{}
%------------------------------------------------------------------
% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
%\pdfoutput=1

@@

% Authors, for the paper (add full first names)
\Author{\hl{Mikhail Laktionov} %MDPI: Please carefully check the accuracy of names and affiliations. 
%MDPI: The name of this author is different from the submitting system. Please confirm which one is correct.
 $^{1, 2}$\orcidA{}, Lucie Nová $^{1}$\orcidB{} and Oleg V. Rud $^{1, 3}$*\orcidC{}}

@@

% MDPI internal command: Authors, for citation in the left column
\AuthorCitation{\hl{Laktionov, M.;} %MDPI: Please check all author names carefully.
 Nová, L.; Rud, O.V.}
% If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad Department of Physical and Macromolecular Chemistry, Faculty of Science, Charles University in Prague, \hl{Czech Republic}%MDPI: please add the city and post code
; miklakt@gmail.com (M.L.); lucie.nova@natur.cuni.cz (L.N.)\\
$^{2}$ \quad St. Petersburg National Research University of Information Technologies, Mechanics and Optics, \hl{St.~Petersburg}%MDPI: please add the post code
, Russia\\
$^{3}$ \quad Institute of Macromolecular Compounds of Russian Academy of Sciences, \hl{Saint-Petersburg}, Russia}

@@

We propose the desalination process as a cascade of gel compressions and decompressions, lowering the  supernatant salinity.
%it is worth to note that we are not limited with this process only.
In this process, the gel is supposed to be compressed/decompressed in an \emph{\hl{open system} %MDPI: Please confirm if the italics should be retained.
} and in a \emph{closed system}.
The compression in the \emph{\hl{open system}} assumes that the gel is in thermodynamic equilibrium with a huge (effectively \textbf{\hl{infinite}}) reservoir of aqueous solution, whereas 
the \emph{\hl{closed system}} implies that the gel is in equilibrium with a \textbf{\hl{finite} %MDPI: Please confirm if the bold should be retained.
} reservoir of an aqueous solution.

@@

	\caption{\hl{Diamond-like network} %MDPI: Please cite the figure in the text and ensure the first citation of each figure appears in numerical order.
 in the simulation box. Color code represents
		the individual ion types (red: $\na$, blue: $\cl$, yellow: $\ca$)
		and the hydrogel (grey: neutral segment ($\AH$), cyan: charged segment
		($\A$)). \label{fig:open and closed}}
\end{figure}
%In the case of \emph{open system} the gel is the grand canonical ensemle, as $\mu, V, T = \mbox{Const}$, while the case of \emph{closed system} the gel and the explicit reservoir are a variation of Gibbs ensemble \cite{Panagiotopoulos1988b} performed with constant volume for the both boxes, equal chemical potentials of ion species at constant temperature.

@@

\includegraphics[width=0.7\textwidth]{{figures/simgibbs}.pdf}\caption{Diamond-like network in the simulation box. Color code represents
the individual ion types (red: $\na$, blue: $\cl$, yellow: $\ca$)
and the hydrogel (gray: neutral segment ($\AH$), cyan: charged segment
($\A$)). \label{fig:diamond}}
\end{figure}
Each monomer unit of the network carries a negative elementary electric charge.
Except for the gel monomers, the monovalent co- and counter-ions, $\cl$ and $\na$, are present in the simulation box.
The total electric charge of all the particles in the box is zero; therefore the number of $\na$ ions exceeds the number of $\cl$ ions by $\Ngel$.

@@

@@

\section{Conclusions}
\textls[-5]{We have modeled compression of a polyelectrolyte gel in thermodynamic equilibrium with a supernatant aqueous solution of limited amount.
We have shown that compression of the gel decreases the supernatant salinity.
We employed this phenomenon to model water desalination.
The desalination was done as a sequential combination of two processes:
(1) swelling of the gel in an \emph{open system}, exchanging ions with a large reservoir at constant salinity;
(2) compression of the gel in a \emph{closed system}, during which the gel exchanges ions with a small reservoir, affecting its salinity.
We estimated the energy consumption needed for producing one liter of potable water from brine and have shown that the proposed gel compression method may compete with modern \hl{desalination~technologies.} %MDPI: we noticed that the supplementary file has been uploaded in submitting system. please confirm if there is a supplementary, if so, please cite it in main text and add section of supplementary materils “\supplementary{The following are available at \linksupplementary{s1}, Figure S1: title, Table S1: title, Video S1: title.} ”. Please note that the reference in supplementary should be also included in mian text.

}

@@

\institutionalreview{\hl{~}}%MDPI: In this section, please add the Institutional Review Board Statement and approval number for studies involving humans or animals. Please note that the Editorial Office might ask you for further information. Please add ``The study was conducted according to the guidelines of the Declaration of Helsinki, and approved by the Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE (protocol code XXX and date of approval).'' OR ``Ethical review and approval were waived for this study, due to REASON (please provide a detailed justification).'' OR ``Not applicable'' for studies not involving humans or animals. You might also choose to exclude this statement if the study did not involve humans or animals.

\informedconsent{\hl{~}}%MDPI: Any research article describing a study involving humans should contain this statement. Please add ``Informed consent was obtained from all subjects involved in the study.'' OR ``Patient consent was waived due to REASON (please provide a detailed justification).'' OR ``Not applicable'' for studies not involving humans. You might also choose to exclude this statement if the study did not involve humans. Written informed consent for publication must be obtained from participating patients who can be identified (including by the patients themselves). Please state ``Written informed consent has been obtained from the patient(s) to publish this paper'' if applicable.

\dataavailability{\hl{~}} %MDPI: In this section, please provide details regarding where data supporting reported results can be found, including links to publicly archived datasets analyzed or generated during the study. Please refer to suggested Data Availability Statements in section ``MDPI Research Data Policies'' at \url{https://www.mdpi.com/ethics}. You might choose to exclude this statement if the study did not report any data.

\acknowledgments{\hl{This research was supported by the Czech Science Foundation (grant 19-17847Y) and the 
government of the Russian Federation (grant number 14.W03.31.0022).} %MDPI: The contents in acknowledgments are the same as those in funding. please confirm if it can be deleted in acknowledgments

}

@@

\bibitem[Miller(2003)]{Miller2003}
Miller, J.
\newblock \emph{Review of Water Resources and Desalination Technologies};
\newblock Technical Report;  \hl{2003}%MDPI: Please add the name of the publisher and the location of it.
.
\newblock
  https://doi.org/10.2172/809106.%}{\detokenize{10.2172/809106}}}.

### 2022-10-14 MDPI/proof/manuscript.v5/temp.tex
proofreading

\begin{table}[H]
\caption{Estimates of  %MDPI: Please confirm if the format of the table should be retained, e.g., the color of the font, the text direction, vertical line
% Oleg Rud: Confirm the colors and the directions of the table must retain
desalination efficiency. All units are calculated per one mol of gel segments. Values in brackets correspond to crossection of `red' and `orange' lines on the plots in Figures \ref{fig: PV and CV} and \ref{fig: NV and CN}.
\label{tab: table}}

>>>>

\begin{table}[H]
\caption{\hl{Estimates of } %MDPI: Please confirm if the format of the table should be retained, e.g., the color of the font, the text direction, vertical line
desalination efficiency. All units are calculated per one mol of gel segments. Values in brackets correspond to crossection of `red' and `orange' lines on the plots in Figures \ref{fig: PV and CV} and \ref{fig: NV and CN}.
\label{tab: table}}
